---
title: "Survival Analysis"
output: html_notebook
---

```{r}
#install.packages(c("survival", "survminer"))
library(magrittr)
library(tidyr)
library(VIM)
library(tidyverse)
library(ggplot2)
library(readxl)
library(readr)
library("survival")
library("survminer")

```

# http://www.sthda.com/english/wiki/survival-analysis-basics

# https://www.youtube.com/watch?v=qt2ufTPCWwI



```{r}
# Load the data
df <- read.csv("new_variables_df.csv")
```

# top 5 rows
```{r}
head(df)
```

# Histogram of survival
```{r}
hist(df$timetoeventcvddate, xlab = "Length of Event Time", main = "Histogram of Event Time in Offspring cohort")
```
We see that the event time are slightly skewed due to the fact that events tend to happen at a latter date in life and the offspring cohort had a accesibility of healthcare and more informed than the parent cohort. 



```{r}
data12 <- df[complete.cases(df), ]
fit <- survfit(Surv(timetoeventcvddate, cvd) ~ sex, data = data12)
print(fit)
```
This tells us that for the 1758 male in the offpsring cohort, 716 were uncensored (followed for the entire time, until occurance of an event). Of the 1828 women in the offspring cohort 496 were uncensored. Male have a median of 82 years survival time and Female have a median of 88 years survival time. The 95% confidence interval for the median survival time for 716 uncensored male is (80-83) whereas for female is (87-91) 


```{r}
summary(survfit(Surv(timetoeventcvddate, cvd) ~ sex, data = data12))
```


```{r}
data12 = df %>% mutate(alload_cat =
                     case_when(alload <= 3 ~ 0, 
                                alload >3 & alload <=7  ~ 1,
                                alload >7 & alload <= 10  ~ 2,
                                
                               ))
```

# Compute survival curves: survfit()

```{r}
data12 <- df[complete.cases(df), ]
fit <- survfit(Surv(timetoeventcvddate, cvd) ~ alload+sex, data = data12)
print(fit)
```

```{r}
data11 <- df[ , c("sex","cvd", "alload",  "timetoeventcvddate")]

head(data11)
```
```{r}
 data11 <- data11[complete.cases(data11), ]
```



# Compute survival curves: survfit()

```{r}
fit2 <- survfit(Surv(timetoeventcvddate, cvd) ~ sex +alload, data = data11)
print(fit2)
```

```{r}
# Summary of survival curves
summary(fit2)
# Access to the sort summary table
summary(fit2)$table
```


```{r}
d <- data.frame(time = fit$time,
                  n.risk = fit$n.risk,
                  n.event = fit$n.event,
                  n.censor = fit$n.censor,
                  surv = fit$surv,
                  upper = fit$upper,
                  lower = fit$lower
                  )
head(d)
```



```{r}
# Change color, linetype by strata, risk.table color by strata
ggsurvplot(fit,title="Survival by Sex in 'Cardiovascular' allostatic load", xlab = "Follow Up Time (Years)",
          pval = TRUE,
          pval.method=FALSE, 
          conf.int = FALSE,
          risk.table = TRUE, # Add risk table
          risk.table.pos="out",
          risk.table.y.text.col=FALSE,
          risk.table.col = "strata",# Change risk table color by groups
          #legend.title="Sex",
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_classic2(base_size=8, base_family = "Arial"),
           font.family = "Arial", # Change ggplot2 theme
          palette = c("red", "blue", "yellow", "green", "orange", "purple")
          )

```




```{r}
fit2 <- survfit(Surv(timetoeventcvddate, cvd) ~ sex +alload, data = data12)

p <- ggsurvplot(fit2, title="Survival by Sex in 'Cardiovascular' Data", xlab = "Follow Up Time (Years)",
  conf.int = FALSE, 
  pval=TRUE, 
  pval.method=TRUE, 
  risk.table=TRUE,
  risk.table.pos="out",
  risk.table.col="black",
  risk.table.y.text.col=FALSE,
  tables.theme = theme_cleantable(),
  #legend.labs=c("Male","Female"),
  font.tickslab = c(8),
  legend.title="Sex",
  ggtheme = theme_classic2(base_size=8, base_family = "Arial"),
           font.family = "Arial"
)
ggpar(p, 
      font.main = c(8),
      font.x = c(8),
      font.y = c(8),
      font.caption = c(8), 
      font.legend = c(8), 
      font.tickslab = c(8))
```

# Cox-PH model

```{r}
colnames(data12)
```


```{r}
cox.mod <- coxph( Surv(timetoeventcvddate, cvd) ~sex + age1+ bmi_imp +years_of_education + sbp+ dbp+ alb + s_creatinine_imp+creat+crp_imp+tchol+hba1c+hdlchol+hdlchol_ratio+educ+maritalst, data = data12)
```

```{r}
summary(cox.mod)
```

# Only use a subset of columns
```{r}

```










# Using Cox Proportional Hazard Model 
```{r}
cph = CoxPHFitter()
cph.fit
```


#Count the cvd events for males and Females

```{r}

 table(data12$cvd, data12$sex)
#offcvd_events<- filter(data12, cvd==1)
#head(offcvd_events,2)
```


# Compute survival curves: survfit()

```{r}
fit2 <- survfit(Surv(timetoeventcvddate, cvd) ~ sex, data = data12)
print(fit2)
```


```{r}
d2 <- data.frame(time2 = fit2$time,
                  n.risk2 = fit2$n.risk,
                  n.event2 = fit2$n.event,
                  n.censor2 = fit2$n.censor,
                  surv2 = fit2$surv,
                  upper2 = fit2$upper,
                  lower2 = fit2$lower
                  )
head(d2)
```


```{r}
# Change color, linetype by strata, risk.table color by strata
ggsurvplot(fit2,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"))
```

```{r}
ggsurvplot(
   fit2,                     # survfit object with calculated statistics.
   pval = TRUE,             # show p-value of log-rank test.
   conf.int = TRUE,         # show confidence intervals for 
                            # point estimaes of survival curves.
   conf.int.style = "step",  # customize style of confidence intervals
   xlab = "Time in days",   # customize X axis label.
   break.time.by = 100,     # break X axis in time intervals by 200.
   ggtheme = theme_light(), # customize plot and risk table with a theme.
   risk.table = "abs_pct",  # absolute number and percentage at risk.
  risk.table.y.text.col = T,# colour risk table text annotations.
  risk.table.y.text = FALSE,# show bars instead of names in text annotations
                            # in legend of risk table.
  ncensor.plot = TRUE,      # plot the number of censored subjects at time t
  surv.median.line = "hv",  # add the median survival pointer.
  legend.labs = 
    c("Male", "Female"),    # change legend labels.
  palette = 
    c("#E7B800", "#2E9FDF") # custom color palettes.
)
```

```{r}
ggsurvplot(fit2,
          conf.int = TRUE,
          risk.table.col = "strata", # Change risk table color by groups
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"),
          xlim = c(0, 600))
```

```{r}
ggsurvplot(fit2,
          conf.int = TRUE,
          risk.table.col = "strata", # Change risk table color by groups
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"),
          fun = "event")
```

```{r}
ggsurvplot(fit2,
          conf.int = TRUE,
          risk.table.col = "strata", # Change risk table color by groups
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"),
          fun = "cumhaz")
```

```{r}
res.sum <- surv_summary(fit2)
head(res.sum)
```

```{r}
require("survival")
fit3 <- survfit( Surv(timetoeventcvddate, cvd) ~ sex + bmi_imp + hba1c,
                data = data12 )
```


```{r}
# Plot survival curves by sex and facet by bmi and spb
ggsurv <- ggsurvplot(fit3, fun = "event", conf.int = TRUE,
                     ggtheme = theme_bw())
   
ggsurv$plot +theme_bw() + 
  theme (legend.position = "right")+
  facet_grid(bmi_imp ~ hba1c)
```

```{r}
ggsurvplot(fit2, surv.median.line = "hv")
```


```{r}
ggsurvplot(fit2, conf.int=TRUE, pval=TRUE, risk.table=TRUE, 
           legend.labs=c("Male", "Female"), legend.title="Sex",  
           palette=c("dodgerblue2", "orchid2"), 
           title="Kaplan-Meier Curve for Cardiovsacular Survival", 
           risk.table.height=.25)
```

```{r}
ggsurvplot(survfit(Surv(timetoeventcvddate, cvd)~hba1c_cat, data=offcvd_events), pval=TRUE,  surv.median.line = "hv")
```

```{r}


ggsurvplot(survfit(Surv(timetoeventcvddate, cvd)~alb_cat, data=offcvd_events), pval=TRUE,  surv.median.line = "hv")
```

# Kaplan-Meier estimate


```{r}
# Estimate the survivor function pretending that all censored observations are actual observations.
km_wrong <- survfit(Surv(timetoeventcvddate) ~ 1, data = offcvd_events)
```


```{r}
# Estimate the survivor function from this dataset via kaplan-meier
km <- survfit(Surv(timetoeventcvddate, cvd) ~1, data=offcvd_events)
```

```{r}
ggsurvplot_combine(list(correct = km, wrong = km_wrong))
```
```{r}
ggsurvplot(km)

# Add the risk table to plot
# ggsurvplot(km, risk.table = TRUE)

# Add a line showing the median survival time
ggsurvplot(km, risk.table = TRUE, surv.median.line = "hv")
```

```{r}
# Weibull model
wb <- survreg(Surv(timetoeventcvddate, cvd) ~ 1, data = offcvd_events)

# Compute the median survival from the model
predict(wb, type = "quantile", p = 1-0.5, newdata = data.frame(1))
```

Half the patients live longer than 65 years and half die before.

```{r}
# 70 Percent of patients survive beyond time point...
predict(wb, type = "quantile", p = 1-0.7, newdata = data.frame(1))
```

70 out of 100 people survive more than 57.5 years.

```{r}
# Retrieve survival curve from model probabilities 
surv <- seq(.99, .01, by = -.01)

# Get time for each probability
t <- predict(wb, type = "quantile", p = 1-surv, newdata = data.frame(1))

# Create data frame with the information
surv_wb <- data.frame(time = t, surv = surv)

# Look at first few lines of the result
head(data.frame(time=t, surv=surv))
```

Survival curve goes down with time.


```{r}
# Create data frame with the information needed for ggsurvplot_df
surv_wb <- data.frame(time = t, surv = surv, 
  upper = NA, lower = NA, std.err = NA)

# Plot
ggsurvplot_df(fit = surv_wb, surv.geom = geom_line)
```

```{r}
# Estimate a Weibull model
wbmod <- survreg(Surv(timetoeventcvddate, cvd) ~ sex, data = offcvd_events)
coef(wbmod)
```


```{r}
# Weibull model
wbmod <- survreg(Surv(timetoeventcvddate,  cvd) ~ alb_cat + hba1c_cat, data = offcvd_events)

# Imaginary patients
newdat <- expand.grid(
  albumin = levels(offcvd_events$alb_cat),
  hba1c = quantile(offcvd_events$bmi, probs = c(0.25, 0.5, 0.75)))
newdat
```

```{r}
# Compute survival curves
surv <- seq(.99, .01, by = -.01)
t <- predict(wbmod, type = "quantile", p = 1-surv,
  newdata = newdat)
```

```{r}
# Use cbind() to combine the information in newdat with t
surv_wbmod_wide <- cbind(newdat, t)
  
# Use melt() to bring the data.frame to long format
surv_wbmod <- melt(surv_wbmod_wide, id.vars = c("horTh", "tsize"), variable.name = "surv_id", value.name = "time")
```

```{r}
# Use surv_wbmod$surv_id to add the correct survival probabilities surv
surv_wbmod$surv <- surv[as.numeric(surv_wbmod$surv_id)]
```

```{r}
# Add columns upper, lower, std.err, and strata to the data.frame
surv_wbmod[, c("upper", "lower", "std.err", "strata")] <- NA

# Plot the survival curves
ggsurvplot_df(surv_wbmod, surv.geom = geom_line,
  linetype = "horTh", color = "tsize", legend.title = NULL)
```



```{r}
# Weibull model
wbmod <- survreg(Surv(timetoeventcvddate, cvd) ~ educ, data = offcvd_events)

# Log-Normal model
lnmod <- survreg(Surv(timetoeventcvddate, cvd) ~ educ, data = offcvd_events, dist = "lognormal")

# Newdata
newdat <- data.frame(education = levels(offcvd_events$educ))

# Surv
surv <- seq(.99, .01, by = -.01)

# Survival curve from Weibull model and log-normal model
wbt <- predict(wbmod, type = "quantile", p = 1-surv, newdata = newdat)
lnt <- predict(lnmod, type = "quantile", p = 1-surv, newdata = newdat)
```


```{r}
# Melt the data.frame into long format.
surv_long <- melt(surv_wide, id.vars = c("horTh", "dist"), variable.name = "surv_id", value.name = "time")

# Add column for the survival probabilities
surv_long$surv <- surv[as.numeric(surv_long$surv_id)]

# Add columns upper, lower, std.err, and strata contianing NA values
surv_long[, c("upper", "lower", "std.err", "strata")] <- NA
```



```{r}
# Plot the survival curves
ggsurvplot_df(surv_long, surv.geom =geom_line,
              linetype = "horTh", color = "dist", legend.title = NULL)
```



