---
title: "Data Imputation"
output: html_notebook
---

#import the libraries

```{r}
library(magrittr)
library(tidyr)
library(VIM)
library(tidyverse)
library(ggplot2)
library(readxl)
```

# Load the merged dataset
```{r}

combined_offspring <- read_excel("~/Library/CloudStorage/Box-Box/FHS DATA_Mogos_James/PETER_FHSproject/Project work/data/Offspring cohort/Combined file excel_offspring2.xls")
View(combined_offspring)
```

# check all colums for null values 
```{r}
combined_offspring %>% is.na() %>% colSums()
```
```{r}
head(combined_offspring)
```

```{r}
combined_offspring %>% 
  select(pid, idtype, sex, age1, bmi, a9, a10, a31, a35, a50, a53, a54, a55, a56, crp6, dheas, epi, aldo) %>% 
  aggr(combined= TRUE, numbers=TRUE)

```

```{r}
mean(combined_offspring$age1)
```
```{r}
# Save the data in two different vector
women_weight <- genderweight %>%
  filter(group == "F") %>%
  pull(weight)
men_weight <- genderweight %>%
  filter(group == "M") %>%
  pull(weight)
# Compute t-test
res <- t.test(women_weight, men_weight)
res
# Compute t-test
res <- t.test(wom, men_weight, var.equal = TRUE)
res

```

```{r}
genderweight %>%
  group_by(group) %>%
  get_summary_stats(weight, type = "mean_sd")
```

## Mean Imputation 

# mean imputation of height Weigh in Nhanes dataset

```{r}
nhanes <- NHANES %>% 
  mutate(height_imp = ifelse(is.na(Height), TRUE, FALSE)) %>% 
  mutate(weight_imp = ifelse(is.na(Weight), TRUE, FALSE))



## Replacing the missing values in height and Weight with their respective means

nhanes_imp <- nhanes %>% 
  mutate(height_imp = ifelse(is.na(Height), mean(Height, na.rm = TRUE), Height)) %>% 
  mutate(weight_imp = ifelse(is.na(Weight), mean(Weight, na.rm = TRUE), Weight))
  


nhanes_imp %>% 
  select(Weight, Height, weight_imp, height_imp ) %>% 
  head()
```


# hot-deck imputation - feed forward

```{r}

```

# k-Nearest-Neighbors imputation

```{r}
# sort variables 
vars_by_NAs <- nhanes %>% 
  is.na() %>% 
  colSums() %>% 
  sort(decreasing = FALSE) %>% 
  names()


# Having sorted the variables

nhanes_imp <- nhanes %>% 
  select(vars_by_NAs) %>% 
  kNN(k=5)


```


```{r}

```




# datatpes
```{r}
str(combined_offspring, 10)
```

# Impute the aldo variable 

K=30
```{r}
aldo_epi <- kNN(combined_offspring, k = 5, variable = c("aldo", "epi"))
```

#head

```{r}
head(aldo_epi)
```
# Weighted donors 

```{r}
aldo_epi <- combined_offspring %>% 
  kNN(variable = c("aldo", "epi"),
      k = 5,
      numFun = weighted.mean(),
      weightDist = TRUE)                
```



# Draw a margin plot of bmi vs aldosterone

```{r}
aldo_imp30 %>% 
  select(bmi, aldo, aldo_imp) %>% 
  marginplot(delimiter = "imp", main = "k=30")
```

K15
```{r}
aldo_imp15 <- kNN(combined_offspring, k = 15, variable = "aldo")
```

```{r}
aldo_imp15 %>% 
  select(bmi, aldo, aldo_imp) %>% 
  marginplot(delimiter = "imp", main = "k=15")
```


```{r}
aldo_imp <- kNN(combined_offspring, k = 5, variable = "aldo")
```

# Draw a margin plot of bmi vs aldosterone

```{r}
aldo_imp %>% 
  select(a55, aldo, aldo_imp) %>% 
  marginplot(delimiter = "imp", main = "k=5")
```
# K2

```{r}
aldo_imp2 <- kNN(combined_offspring, k = 2, variable = "aldo")
```


```{r}
aldo_imp2 %>% 
  select(a55, aldo, aldo_imp) %>% 
  marginplot(delimiter = "imp", main = "k=2")
```
# Impute using distance weighted mean

```{r}
aldo_weighted_im <- kNN(combined_offspring,
                        k=3,
                        variable = "aldo",
                        numFun = weighted.mean,
                        weightDist = TRUE)
                        
```


```{r}
head(aldo_weighted_im)
```

```{r}
mean
```


#marginplot

```{r}
aldo_weighted_im %>% 
  select(a138, aldo, aldo_imp) %>% 
  marginplot(delimiter = "imp")
```

```{r}
colnames(combined_offspring)
```

# Subset Data for only the required variables 

```{r}
allostatic_loads <- combined_offspring %>% 
  select(pid, idtype, sex, bmi, a9, a10, a31, a35, a50, a55, a56, crp6, dheas, epi, aldo)
```



# Get combine offsprings Sorted by number of  NAs

```{r}
vars_by_NAs <- combined_offspring %>% 
  is.na() %>% 
  colSums() %>% 
  sort(decreasing = FALSE) %>% 
  names()
```

# Sort combined offsprings variables and feed it to kNN imputation
```{r}
combined_offspring_imp <- combined_offspring %>% 
  select(vars_by_NAs) %>% 
  kNN(variable = c("bmi","a9","a10","a31","a35","a50","a55", "crp6","dheas","epi", "aldo"))
```

# marginplot
```{r}
combined_offspring_imp %>% 
  select(bmi, aldo, aldo_imp) %>% 
  marginplot(delimiter = "imp")
```



#Check 
```{r}
head(combined_offspring_imp)
```

null values
```{r}
combined_offspring_imp %>% is.na() %>% colSums()
```

# Save to xls file
```{r}
write.csv(combined_offspring_imp, "~/Library/CloudStorage/Box-Box/FHS DATA_Mogos_James/PETER_FHSproject/Project work/data/Offspring cohort/imputed_data.csv")
```

#Imputed Dataframe
```{r}
imputed_data <- read_csv("~/Library/CloudStorage/Box-Box/FHS DATA_Mogos_James/PETER_FHSproject/Project work/data/Offspring cohort/imputed_data.csv")
head(imputed_data)
```


```{r}
dim(imputed_data)
```

# Saved as excel
```{r}
library(writexl)
write_xlsx(combined_offspring_imp, "~/Library/CloudStorage/Box-Box/FHS DATA_Mogos_James/PETER_FHSproject/Project work/data/Offspring cohort/imputed_excel.xlsx")
```


# kNN imputation
```{r}
offspring_imp <- kNN(combined_offspring, k = 5, variable = c("crp6", "aldo"))
```

# Looking at the imputed data

```{r}
offspring_imp %>% is.na() %>% colSums()
```
# Model-based imoputation Approach
## linear regression imputation of continues  variables
```{r}
#install.packages('simputation', dependencies=TRUE)
library(simputation)
```


# Single linear regression imputation

```{r}
lin_regression_imp <- impute_lm(combined_offspring, aldo + a55 ~.)
```


# Initialize the missing Values with hoddeck and save missing locations
```{r}
offsprings_m_imp <- hotdeck(combined_offspring)

missing_crp6 <- offsprings_m_imp$crp6_imp
missing_aldo <- offsprings_m_imp$aldo_imp
```

# Iterate over the crp6 and aldo 5 times, imputing them at the originally missing locations

```{r}
for (i in 1:5){
  #prev_iter <- offsprings_m_imp
  offsprings_m_imp$crp6[missing_crp6] <-NA
  offsprings_m_imp <- impute_lm(offsprings_m_imp, crp6 ~ bmi+a9 +a10+a55)
  offsprings_m_imp$aldo[missing_aldo] <-NA
  offsprings_m_imp <- impute_lm(offsprings_m_imp, aldo ~ bmi + a9 +a10 +crp6)
  #diff_crp6 <- c(diff_crp6, mapc(prev_iter$crp6, offsprings_m_imp$crp6))
  #diff_aldo <- c(diff_aldo, mapc(prev_iter$aldo, offsprings_m_imp$aldo))
  
}
```


```{r}
mapc <- function(a, b){
  mean(abs(b - a) / a, na.rm = TRUE)
}
```


# Detecting convergence

```{r}
#library(purrr)
# Detecting the convergence
diff_crp6 <- c()
diff_aldo <- c()
# For loop
for (i in 1:5){
  prev_iter <- offsprings_m_imp
  offsprings_m_imp$crp6[missing_crp6] <-NA
  offsprings_m_imp <- impute_lm(offsprings_m_imp, crp6 ~ bmi+a9 +a10+a55)
  offsprings_m_imp$aldo[missing_aldo] <-NA
  offsprings_m_imp <- impute_lm(offsprings_m_imp, aldo ~ bmi + a9 +a10 +crp6)
  diff_crp6 <- c(diff_crp6, mapc(prev_iter$crp6, offsprings_m_imp$crp6))
  diff_aldo <- c(diff_aldo, mapc(prev_iter$aldo, offsprings_m_imp$aldo))
  
}
```


# Difference between 

```{r}
diff_crp6
```




Load combine datasets
```{r}
model_imp <- impute_lm(combined_offspring, crp6 + aldo ~.) 
```



```{r}
# https://dept.stat.lsa.umich.edu/~jerrick/courses/stat701/notes/mi.html

# https://stefvanbuuren.name/fimd/sec-knowledge.html


#data <- boys[, c("age", "hgt", "wgt", "hc", "reg")]

#data
```

