---
title: "Variables_transformations"
output:
  pdf_document: default
  html_notebook: default
---
# Libraries 


```{r}
library(magrittr)
library(tidyr)
library(VIM)
library(tidyverse)
library(ggplot2)
library(readxl)
```

# Load the dataset

```{r}
imputed_df <- read_csv("~/Library/CloudStorage/Box-Box/FHS DATA_Mogos_James/PETER_FHSproject/Project work/data/Offspring cohort/imputed_data.csv")
#View(imputed_df)
```

```{r}
View(imputed_df)
```

# top 5 

```{r}
head(imputed_df)
```

# generating fasting glucose, defined  if fasting more than 6 hours (g3a453) and glucose levels (g3a715)

```{r}
head(imputed_df$a31)
```

# convert fpg to a1c using the following formular used for eAG (note eAG and FPG are highly correlated) A1C(%) = (Estimated average glucose(mg/dL) +46.7) / 28.7
# https://ebmcalc.com/GlycemicAssessment.htm

```{r}

```

# defining education: G3A491 WHAT IS THE HIGHEST DEGREE OR LEVEL OF SCHOOL YOU HAVE COMPLETED? 0 = NO SCHOOLING1 = GRADES 1-82 = GRADES 9-113 = COMPLETED HIGH SCHOOL (12TH GRADE) OR GED4 = SOME COLLEGE BUT NO DEGREE5 =TECHNICAL SCHOOL CERTIFICATE6 =ASSOCIATE DEGREE (JUNIOR COLLEGE AA,AS)7 =BACHELOR'S DEGREE (BA,AB,BS)8 = GRADUATE OR PROFESSIONAL DEGREE (MASTER'S, DOCTORATE, MD, ETC.). = PREFER NOT TO ANSWER/UNKNOWN (16)

```{r}

```

# reclassify education the following categories:1 elementary 2 high school; 3 some college/asscdegree, 4 college degree (bachelor/graduate)

```{r}
imputed_df <-imputed_df %>% 
  mutate(BMI_categories = 
           case_when(education_classification = "elementary" ~ "1",
                     education_classification = "high school"  ~ "2",
                     education_classification = "some college/asscdegree" ~ "3",
                     education_classification = "college degree" ~ "4"
                     
                     ))
```

# generating income variable from G3A497PLEASE SELECT WHICH INCOME GROUP BEST REPRESENTS YOUR COMBINEDFAMILY INCOME FOR THE PAST 12 MONTHS.1 = UNDER $12,0002 = $ 12,000 -$ 24,0003 = $ 25,000 -$ 49,9994 = $ 50,000 -$ 74,9995 =$ 75,000 -$ 100,0006 =OVER $ 100,000. = PREFER NOT TO ANSWER/UNKNOWN (196)

```{r}
imputed_df <-imputed_df %>% 
  mutate(BMI_categories = 
           case_when(G3A497 >= 100,0002 & G3A497 <=149 ~ "7",
                     G3A497 >= 120 & G3A497 <=149 ~ "0.5",
                     G3A497 >= 120 & G3A497 <=149 ~ "0.5",
                     G3A497 >= 120 & G3A497 <=149 ~ "0.5",
                     G3A497 >= 12,000 & G3A497 <=149 ~ "0.5",
                     G3A497 < 12,000 ~ "1"
                     
                     ))
```


# generating marital status from var g3a482

```{r}

```


# generating bmi from weight in pounds (g3a444) and height inches code=G3A446

```{r}

```

# convert weight lbs to wt in kgs=lbs/2.2046

```{r}

```

# convert ht inches to ht meters=inches*0.0254

```{r}

```

# calculate bmi by formula=BMI = kg/m2

```{r}

```

# first gen ht_m squared

```{r}

```

# generating blood pressure-start with SBP from combination of average of three variables 1, 2nd and 3rd reading of SBP (G3A290, 347,456) and then three readings of DBP

```{r}
imputed_df$sbp <- rowMeans(imputed_df[, c("a53","a55","a57")], na.rm = TRUE)
imputed_df$dbp <- rowMeans(imputed_df[, c("a54","a56","a58")], na.rm = TRUE)
```

```{r}

imputed_df$dbp <- rowMeans(imputed_df[, c("a54","a56","a58")], na.rm = TRUE)
```


```{r}
#sum(imputed_df$a53, na.rm = TRUE)

dim(imputed_df)
```

# Total N of a55

```{r}
sapply(imputed_df$a55, na.rm = TRUE)
```
```{r}
summary(imputed_df)
```


# Generate systolic blood pressure (sbp) 1=high=≥150 0.5=mod= 120 to 149 0=low=<120 

```{r}
imputed_df <-imputed_df %>% 
  mutate(BMI_categories = 
           case_when(a55 >= 150 ~ "1",
                     a55 >= 120 & a55 <=149 ~ "0.5",
                     a55 < 120 ~ "0"
                     
                     ))
```

# Generate diastolic blood pressure dbp 1=high=≥90 0.5=mod= 80 to 89 0=low=<120 ≥80 ?????????

```{r}
imputed_df <-imputed_df %>% 
  mutate( sbp_cat = 
           case_when(a56 >= 90 ~ "1",
                     a56 >= 80 & a56 <= 89 ~ "0.5",
                     a56 < 80 ~ "0"
                     
                     ))
```

# Creating tchol and hdlchol variables

```{r}


imputed_df$tchol <- imputed_df$a9

imputed_df$hdlchol <- imputed_df$a10


```


```{r}
imputed_df$thdlchol_ratio <- imputed_df$tchol/ imputed_df$hdlchol
```

```{r}
#range(imputed_df$thdlchol_ratio)
colnames(imputed_df)
```

# generating HDL cholesterol 1=high<40 0.5=40-59 low risk>60

```{r}
imputed_df <-imputed_df %>% 
  mutate(hdl_cat = 
           case_when(hdlchol < 40 ~ "1", 
                     hdlchol >= 40 & a10 <=59 ~ "0.5",
                     hdlchol > 60 ~ "0"
                     
                     ))
```


# generating total cholesterol 1=high≥240 0.5=200 to 239 low risk=<200

```{r}
imputed_df <-imputed_df %>% 
  mutate(tchol_cat = 
           case_when(tchol >=240 ~ "1",
                    tchol >= 200 &  tchol < 239 ~ "0.5",
                      tchol < 200 ~ "0"
                     
                     ))
```



# generating tchol/HDL ratio cholesterol 1=high>=6 0.5=mod= 5 to <6 0=low risk<5  

```{r}
imputed_df <-imputed_df %>% 
  mutate(thdlchol_ratio = 
           case_when(a9 >=6 ~ "1",
                     a9 >= 5 & a9 <6 ~ "0.5",
                     a9 < 5 ~ "0"
                     
                     ))
```

# generating BMI_categories 1=high=≥30=poor kg/m2, 0.5=25–<30 kg/m2, 0= <25 kg/m2

```{r}

 #thdlchol_ratio= tchol/hdlchol
imputed_df2 <-imputed_df %>% 
  mutate(bmi_cat = 
           case_when(bmi >=30 ~ "1",
                     bmi >= 25 & bmi <30 ~ "0.5",
                     bmi < 25 ~ "0"
                     
                     ))

#1=high=≥30=poor kg/m2

#0.5=25–<30 kg/m2,
#0= <25 kg/m2
```


```{r}
imputed_df2[c(150, 215)]
```
```{r}
mean(imputed_df2$bmi)
```
```{r}
freq <- table(imputed_df2$bmi_cat)
```

```{r}
print(freq)
```

